---
title: "conceptual model DCNP"
output: html_document
date: '2022-08-11'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dagitty)
```

```{r}
source('./R/dag1.R')

g <- dagitty(dag1)

names(g)

plot(g)
```


Let's start with a population of 10,000 people, who have a baseline hazard of death of 2% annually (probability of death for men at 70 is 2.2%, at 75 is 3.5%, at 80 is 5.6%, at 85 is 9.5%, at 90 is 16.2% at 95 is 25.9% https://www.ssa.gov/oact/STATS/table4c6.html )
```{r}
p_death_year <- 0.02

p_death_month <- 1 - (1-p_death_year)^(1/12)




```

Markov model:
You can be healthy, transition to different conditions (e.g., acute cardiac, cancer, pneumonia,  ), and can be treated or not, being in each of the states affect the rate of death. Have a pre--COVID model, then have COVID as a modifier of the rates

compartment 1: healthy

Compartment 2a: Chronic condition in need of monitoring, in control
Compartment 2b: Chronic condition in need of monitoring, out of control

Compartment 3A: Hospitalized acute infectious issue (e.g., flu, pneumonia, NON COVID)

Compartment 3B: severe COVID


Compartment 4A: Depression other mental health (which increase risk for suicide)
Compartment 4B: Untreated severe depression

Compartment 5: DEATH

3A -> 5 transition higher during disruptions?
2A -> 2B transition higher
4A->4B transition higher

3B doesn't exist without COVID; 3B ->5 transition worse during disruption? less bad over time? 


```{r}



```

