---
title: "conceptual model DCNP"
output: html_document
date: '2022-08-11'
---


 
 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dagitty)
```

```{r}
source('./R/dag1.R')

g <- dagitty(dag1)

names(g)

plot(g)
```


Let's start with a population of 10,000 people, who have a baseline hazard of death of 2% annually (probability of death for men at 70 is 2.2%, at 75 is 3.5%, at 80 is 5.6%, at 85 is 9.5%, at 90 is 16.2% at 95 is 25.9% https://www.ssa.gov/oact/STATS/table4c6.html )
```{r}
p_death_year <- 0.02

p_death_month <- 1 - (1-p_death_year)^(1/12)


```

Markov model:
You can be healthy, transition to different conditions (e.g., acute cardiac, cancer, pneumonia,  ), and can be treated or not, being in each of the states affect the rate of death. Have a pre--COVID model, then have COVID as a modifier of the rates

A compartment 1: healthy (No chronic conditions)

B Compartment 2a: Chronic condition in need of monitoring, in control
C Compartment 2b: Chronic condition in need of monitoring, out of control

D Compartment 3A: Hospitalized acute infectious issue (e.g., flu, pneumonia, NON COVID)

E Compartment 3B: severe COVID


F Compartment 4A: Treated depression other mental health (which increase risk for suicide)

G Compartment 4B: Untreated severe depression

Z Compartment 5: DEATH

(To assign):
Acute health event (e.g., MI, stroke)


Transition probabilities (age specific)
p1 A,B,F -> Z (baseline death rates not acute infections or acute chronic conditions or mental)
p2 B -> C (loss of control of chronic conditions)
p3 C -> B (regain control of chronic conditions)
p4 C -> Z (death from uncontrolled chronic conditions)

p5 A,B,C,F,G -> D (probably separate out for each)

p6 A,B,C,F,G -> E (probably separate out for each)

p7 A -> F
p8 A -> G
p9 F -> G

p10 G -> Z
p11 A,B,F -> Z

Many of the probabilities should be a function of COVID/ hospital rates, etc.

3A -> 5 transition higher during disruptions?
2A -> 2B transition higher
4A->4B transition higher

3B doesn't exist without COVID; 3B ->5 transition worse during disruption? less bad over time? 

some possible transitions




```{r}

#Probability of seeking care given emergency situation
## probability of surviving emergency situation in hospital, delayed care

#Probability of death given hospitalization for emergency condition

##Probability of getting elective surgery or unnecessary care
###probability of lethal complications given elective surgery

#probability of death with in-control chronic condition
#probability of death with out-of-control chronic conditon

##Probability of detecting cancer/diabetes/hypertension if screened
###Probability of seeking screening
####Probability of death with or without screening

#Probability of suicide attempt/succesful suicide

#Probability of getting killed by violent crime


#Probability of hospitalization/death due to a non-COVID virus

##Probability of hospitalization due to covid
###Probability of death due to COVID if hospitalized/avoided hospital

```

## Blood pressure

less frequent monitoring associated with loss of control; more frequent monitoring associated with regaining control

frequency of monitoring is time varying and depends on pandemic, and maybe on state of the outcome (check about this)..Multistate model (might not be Markov model because outcome might depend on previous states)

some refs:
ahajournals.org/doi/10.1161/JAHA.120.016174


first a simple 2 state model, where monitoring is uniform
```{r}
t <- 365*10

N <-100000

P_1_2 <- 1/(365*4)
  
P_2_1 <- 1/(365/2)
  
RR_1_2_freq <- 1.2 # effect of less frequent monitoring on P_1_2

RR_2_1_freq <- 1/0.8 # effect of more frequent monitoring on P_2_1

#State 1: In Control hypertension; State 2: Out of control 

State1 <- rep(NA, t)

State2 <- rep(NA, t)

State1[1] <- 30000
  
State2[1] <- N - State1[1]

for(i in 2:t){
  State1[i] <- State1[i-1] - State1[i-1]*P_1_2 + State2[i-1]*P_2_1
  State2[i] <- State2[i-1] - State2[i-1]*P_2_1 +State1[i-1]*P_1_2
}

plot(State1, type='l')
plot(State2, type='l')
  

```
Have 2 populations with 2 states each:
frequent screen, less frequent; in control, out of control

```{r}

t <- 365*10

N <-100000

P_1_2 <- 1/(365*4)
  
P_2_1 <- 1/(365/2)
  
RR_1_2_freq <- 1.2 # effect of less frequent monitoring on P_1_2

RR_2_1_freq <- 1/0.8 # effect of more frequent monitoring on P_2_1

#State 1: In Control hypertension; State 2: Out of control 

State1_infreq <- rep(NA, t)

State2_infreq <- rep(NA, t)

State1_freq <- rep(NA, t)

State2_freq <- rep(NA, t)

#infrequent monitoring
State1_infreq[1] <- 2000
  
State2_infreq[1] <- 5000

#frequent monitoring
State1_freq[1] <- 10000
  
State2_freq[1] <- N - State1_infreq[1] -State2_infreq[1] -State1_freq[1]

#need to think about the transitions here--if out of control, might shift to more frequent monitoring.

for(i in 2:t){
  State1[i] <- State1[i-1] - State1[i-1]*P_1_2 + State2[i-1]*P_2_1
  
  State2[i] <- State2[i-1] - State2[i-1]*P_2_1 +State1[i-1]*P_1_2
}

```

