---
title: "conceptual model DCNP"
output: html_document
date: '2022-08-11'
---


 
 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dagitty)
```

```{r}
source('./R/dag1.R')

g <- dagitty(dag1)

names(g)

plot(g)
```


Let's start with a population of 10,000 people, who have a baseline hazard of death of 2% annually (probability of death for men at 70 is 2.2%, at 75 is 3.5%, at 80 is 5.6%, at 85 is 9.5%, at 90 is 16.2% at 95 is 25.9% https://www.ssa.gov/oact/STATS/table4c6.html )
```{r}
p_death_year <- 0.02

p_death_month <- 1 - (1-p_death_year)^(1/12)


```

Markov model:
You can be healthy, transition to different conditions (e.g., acute cardiac, cancer, pneumonia,  ), and can be treated or not, being in each of the states affect the rate of death. Have a pre--COVID model, then have COVID as a modifier of the rates

A compartment 1: healthy (No chronic conditions)

B Compartment 2a: Chronic condition in need of monitoring, in control
C Compartment 2b: Chronic condition in need of monitoring, out of control

D Compartment 3A: Hospitalized acute infectious issue (e.g., flu, pneumonia, NON COVID)

E Compartment 3B: severe COVID


F Compartment 4A: Treated depression other mental health (which increase risk for suicide)

G Compartment 4B: Untreated severe depression

Z Compartment 5: DEATH

(To assign):
Acute health event (e.g., MI, stroke)


Transition probabilities (age specific)
p1 A,B,F -> Z (baseline death rates not acute infections or acute chronic conditions or mental)
p2 B -> C (loss of control of chronic conditions)
p3 C -> B (regain control of chronic conditions)
p4 C -> Z (death from uncontrolled chronic conditions)

p5 A,B,C,F,G -> D (probably separate out for each)

p6 A,B,C,F,G -> E (probably separate out for each)

p7 A -> F
p8 A -> G
p9 F -> G

p10 G -> Z
p11 A,B,F -> Z

Many of the probabilities should be a function of COVID/ hospital rates, etc.

3A -> 5 transition higher during disruptions?
2A -> 2B transition higher
4A->4B transition higher

3B doesn't exist without COVID; 3B ->5 transition worse during disruption? less bad over time? 

some possible transitions




```{r}

#Probability of seeking care given emergency situation
## probability of surviving emergency situation in hospital, delayed care

#Probability of death given hospitalization for emergency condition

##Probability of getting elective surgery or unnecessary care
###probability of lethal complications given elective surgery

#probability of death with in-control chronic condition
#probability of death with out-of-control chronic conditon

##Probability of detecting cancer/diabetes/hypertension if screened
###Probability of seeking screening
####Probability of death with or without screening

#Probability of suicide attempt/succesful suicide

#Probability of getting killed by violent crime


#Probability of hospitalization/death due to a non-COVID virus

##Probability of hospitalization due to covid
###Probability of death due to COVID if hospitalized/avoided hospital

```

## Blood pressure

less frequent monitoring associated with loss of control; more frequent monitoring associated with regaining control

frequency of monitoring is time varying and depends on pandemic, and maybe on state of the outcome (check about this)..Multistate model (might not be Markov model because outcome might depend on previous states)

some refs:
ahajournals.org/doi/10.1161/JAHA.120.016174


first a simple 2 state model, where monitoring is uniform
```{r}
t <- 365*10

N <-100000

P_1_2 <- 1/(365*4)
  
P_2_1 <- 1/(365/2)
  
RR_1_2_freq <- 1.2 # effect of less frequent monitoring on P_1_2

RR_2_1_freq <- 1/0.8 # effect of more frequent monitoring on P_2_1

#State 1: In Control hypertension; State 2: Out of control 

State1 <- rep(NA, t)

State2 <- rep(NA, t)

State1[1] <- 30000
  
State2[1] <- N - State1[1]

for(i in 2:t){
  State1[i] <- State1[i-1] - State1[i-1]*P_1_2 + State2[i-1]*P_2_1
  State2[i] <- State2[i-1] - State2[i-1]*P_2_1 +State1[i-1]*P_1_2
}

plot(State1, type='l')
plot(State2, type='l')
  

```


Have 2 populations with 2 states each:
frequent screen, less frequent; in control, out of control

CK: The recommendation for blood pressure measurement changes according to HTN control – those with uncontrolled HTN should be monitored every 4-6 weeks, and those with controlled HTN every 6 months (some would say 3-6). I would think if we make 4 states like you describe there could me movement between any two – e.g., someone has controlled HTN and is monitored frequently [state 1], and then they have a lapse in monitoring and become uncontrolled [state 4].


4 states:
 incontrol_freq_1
 outcontrol_freq_2
 incontrol_infreq_3
 outcontrol_infreq_4
 
 All transitions are possible
 

popsizes: 
Table2 495k people :In control BP: 79808 had readings <2months; 20554 at 3 months, 106028 4-7 months( 91-210 days); 289031 had readings more than 7 months apart; unadjusted RR for >7 months ranged from 1.07-1.12 (RR1); for adjusted: 1.05-1.09, RR for 3-7 was ~1; rr for 0-1months was <1 for developing hypertension (transitions from 3-4)

table 3;  1.1 million people: start out of control, if monitor every 30-210 days (1-7 months) RR=0.88-0.91 to be less likely to reduce BP compared to those that monitor monthly; if >7 months, RR=0.79-0.84 to reduce BP.


Note these Probabilities might not be right

56% of those with controlled HT became uncontrolled at first pandemic reading (P_3_4=0.56)
--P_1_2 = P_3_4*(1/1.10)

20.7% of those with uncontrolled HT became controlled  (P_2_1 = 0.21) 
--P_4_3 = P_2_1*0.8


use approach of Kissler and grad to convert prob to rate and back when calculatin excess

```{r}

t <- 365*500  #100 year simulation period


#Main transitions..cumulative probabilities over a 2 year period. calculate daily prob as 
#1 - exp(log(1-cum_prob)/t)

cum_P_3_4=0.54 #Prob going from in control to out of control with infrequent monitoring

cum_P_1_2 = 1 - (1-cum_P_3_4)^(1/1.10) #Prob going from in control to out of control with frequent monitoring

cum_P_2_1 = 0.21 #Prob going from out of control to in control with frequent monitoring

cum_P_4_3 = 1 - (1-cum_P_2_1)^0.8  #Prob going from out of control to in control with infrequent monitoring

cum_P_1_3 = 0.2 #Prob of in control frequent monitoring to in control infrequent monitoring

cum_P_3_1 = 0.3 #Prob of in control infrequent monitoring to in control frequent monitoring

cum_P_4_2 = 0 #Prob of out of control infrequent monitoring to out of control frequent monitoring

cum_P_4_1 = 0 #Prob of out of control infrequent monitoring to in control frequent monitoring

cum_P_1_4 <- 0 #prob of in control frequent to out of conrol infrequent

cum_P_2_3 <- 0 #prob of out of control frequent to in control, infrequent

cum_P_2_4 <- 0 #prob of out of control frequnt to out of control infrequent

cum_P_3_2 <- 0 #prob of in control infrequnt to out of control frequent.


#Convert cumulative to daily prob:
P_1_2 = 1 - exp(log(1-cum_P_1_2)/(365*2))

P_1_3 = 1 - exp(log(1-cum_P_1_3)/(365*2))

P_1_4 = 1 - exp(log(1-cum_P_1_4)/(365*2))

P_2_1 = 1 - exp(log(1-cum_P_2_1)/(365*2))

P_2_3 = 1 - exp(log(1-cum_P_2_3)/(365*2))
P_2_4 = 1 - exp(log(1-cum_P_2_4)/(365*2))

P_3_1 = 1 - exp(log(1-cum_P_3_1)/(365*2))
P_3_2 = 1 - exp(log(1-cum_P_3_2)/(365*2))
P_3_4 = 1 - exp(log(1-cum_P_3_4)/(365*2))

P_4_1 = 1 - exp(log(1-cum_P_4_1)/(365*2))

P_4_2 = 1 - exp(log(1-cum_P_4_2)/(365*2))

P_4_3 = 1 - exp(log(1-cum_P_4_3)/(365*2))




#Frequent monitoring: State 1: In Control hypertension; State 2: Out of control 

State1 <- rep(NA, t)

State2 <- rep(NA, t)

State3 <- rep(NA, t)

State4 <- rep(NA, t)

#frequent monitoring, in control
State1[1] <- 79808 + 20554

#frequent monitoring, out of control
State2[1] <- 500000


#infrequent monitoring, in control 
State3[1] <- 495000 - State1[1]

#infrequent monitoring, out of control
State4[1] <- 1.1e6 - State2[1]



for(i in 2:t){
  State1[i] <- State1[i-1] - State1[i-1]*(P_1_2 + P_1_3 + P_1_4) + State2[i-1]*P_2_1 + State3[i-1]*(P_3_1) + + State4[i-1]*(P_4_1)
  
  State2[i] <- State2[i-1] - State2[i-1]*(P_2_1+P_2_3+P_2_4) +State1[i-1]*P_1_2 +State3[i-1]*P_3_2 +State4[i-1]*P_4_2

  State3[i] <- State3[i-1] - State3[i-1]*( P_3_1 + P_3_2 + P_3_4 ) +  State1[i-1]*P_1_3 + State2[i-1]*P_2_3 + State4[i-1]*P_4_3 

  State4[i] <- State4[i-1] - State4[i-1]*(P_4_1 + P_4_2 + P_4_3) +State1[i-1]*P_1_4 + +State2[i-1]*P_2_4 + State3[i-1]*P_3_4
  
}

plot(State1[-c(1:50000)])
plot(State2)
plot(State3)
plot(State4)

```

